Gravitational Lens Modeling in a Citizen Science Context

### 2014-07-29 text/ms.tex
filled in some references

\definecolor{purple}{RGB}{150,0,200}
\newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf #1}}
\let\phil\relax

@@

We develop a method to enable collaborative modelling of gravitational
lenses, and lens candidates, by non-professional lens enthusiasts.  It
uses an existing free-form modelling program (GLASS), but
\phil{enables the input to this code to be provided in a novel way,}
via a user-provided diagram that is essentially a sketch of an
arrival-time surface.

An implementation \phil{of this method, \spl} is tested in a modelling
challenge using 29 simulated lenses \phil{drawn from the} set created
for \phil{the \sw citizen-science lens search}.  We find that
\phil{volunteers from this online community} asserted the image
parities and time ordering consistently in some lenses, but made
errors in other lenses depending on the image morphology. \phil{While}
errors in image parity and time ordering lead to  large errors in the
mass distribution, the enclosed mass \phil{was found to be more
robust: the model-derived Einstein radii found by the volunteers were
consistent with those produced by one of the professional team,
suggesting that given the appropriate tools, gravitational lens
modelling is a data analysis activity that can be crowd-sourced to
good effect.} Shortcomings \phil{of the software implementation 
include} (a)~lens profiles that tend to be systematically shallower
than the lenses \phil{(leading to systematic over-estimation of the
Einstein radius  by some $\sim25\%$}, and (b)~lack of detailed
\phil{arc} modelling.  Ideas for improvement are discussed.

@@

Imaging surveys now under way aim to increase the inventory of lenses
another ten or a hundred fold \phil{\citep[see
e.g.][]{Marshall2005,OguriMarshall2010}, with both automated and
visual search techniques proposed
\citep[e.g.][]{Marshall2009,More2012ApJ,Gavazzi2014}. For example, \sw
(Marshall et al, in prep, More et al, in prep) is a citizen science
project\footnote{\tt http://www.spacewarps.org} in which volunteers
are presented sky-survey images and invited to identify lens
candidates, by eye.}  Simulated lenses are mixed in with the data,
both to help train volunteers on what to look for, and to provide
measures of the effectiveness of the search.  The motivation for \sw
is to enable volunteers, some of whom had previously serendipitously
identified lens candidates on earlier citizen-science surveys,
\phil{either} to make discoveries missed in automatic searches by
software robots, \phil{or to perform the necessary inspection of an
automatically-generated sample, for quality control}.  Robots can be
built to be good at detecting clean lensing system in uncrowded fields
with high signal-to-noise, but in more general test situations robots
miss lenses (low completeness) or contaminate the results with
non-lenses (low purity) \citep{Marshall2009}.

\phil{The encouraging early results from the first \sw lens search,
carried out on the  $\simeq172$ square degree Canada-France-Hawaii
Telescope Legacy Survey (CFHTLS) imaging by over 30,000 volunteers
(Marshall et al, in prep, More et al, in prep) prompted the question:}
could the modelling of the lenses also be done by the volunteers?  If
so, modelling could help assess prioritise lens candidates at an early
stage, which would be very useful with new wide-field and sensitive
surveys, which will yield thousands of lens candidates. There are
several software tools for lensing modelling available, and work has
been done on generic interfaces \citep[e.g.][]{2014A&C.....5...28L}. 
Indeed, some early designs for \sw included a prototype lens modelling
tool \citep{2010AAS...21543527N}. Moreover, some \sw volunteers are
quite experienced from earlier projects, having spent a thousand hours
or more with data, and are very interested in more demanding
projects.  The interests of citizen-science communities are just
beginning to be studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but it
is clear that some volunteers welcome open-ended challenges, and
sometimes these have led to new scientific results: one example is the
discovery of an exceptional extra-solar planet
\citep{2013ApJ...768..127S}; another is the development of new
algorithms for protein folding \citep{Khatib22112011}.  All these are
grounds for optimism.  There is, however, a basic difficulty in strong
gravitational lensing. Lensed images do not look much like their
source, and still less do they resemble the lensing-mass
distribution.  To model a lens, one needs either to do a lot of random
guessing, or to have a good intuition for what works.

@@

In \secref{SpaghettiLens} we describe the \spl program, which
implements \phil{the above scheme}.  \spl is an \phil{interface to and
extension of} the GLASS framework for modelling lenses
\citep{2014arXiv1401.7990C}.  We will not go into software details in
this paper, but instead concentrate on lens modelling per se, but we
remark that \spl is designed to be friendly to the forum style of
citizen-science projects, and enable incremental \phil{collaborative}
model refinement by different people, without sacrificing any of the
technical features of GLASS.

In \secref{mod_challenge} we describe a modelling challenge where a
diverse sample of 29 simulated lenses was modelled multiple times
\phil{by a small number of \sw volunteers} using \spl.  The models
were then examined in two ways.  One was whether the spaghetti diagram
was correct.  The other was the recovery of the Einstein radius of the
lens.  In addition, we show some visual comparisons of the actual and
recovered lens shape and radial profile, and identify some areas to
improve. Profile and shape recovery with GLASS has been studied in
more detail in \citep{2014arXiv1401.7990C}.

@@

\spl is not part of \sw, but it makes use of the \sw infrastructure,
in particular, the image database and the \phil{discussion
forum.\footnote{\texttt{http://talk.spacewarps.org}}  The forum is
essential for establishing contact between interested members of the
\sw community and the project science team, and then for enabling
collaboration between them. We were able to collaborate together on 
modelling objects from \sw in the usual style of medium-sized
astronomical collaborations, with video-conferencing and in-person
meetings where possible. Preliminary results were immediately
summarized on modelling threads on the forum, and anyone interested
was made welcome to join at any time.} 
%
% PJM: The following paragraph is not accurate: the manipulation of 
%  FITS images in the Tools viewer, for example, is done client side.
%  In future I very much hope to see lens modeling enabled in this
%  mode as well!
%
% Another aspect of \sw (and of the Zooniverse family of projects in
% general) that \spl adopts, is to keep all user interaction in a web
% browser, with intensive computing done on a server.  Several
% interesting design and implementation issues arise from these
% considerations, but we will not go into them in this paper,
% concentrating instead on aspects specific to lens modelling.
%
Modelling with \spl involves three stages, markup of the image,
followed by intensive \phil{numerical computation carried out on a
server in the background,} followed by review
of diagnostics and possible discussion.  We now describe each of
these.

@@

The markup tool allows only valid lensing configurations to be
entered.  The user does not need to think explicitly the image
parities (though the markup tool provides this information using
colour codes) or about time-ordering, or worry about the odd-image
theorem.  The exact placement of the loops in a spaghetti diagram has
no significance.  Only the hierarchy of which loop is inside which is
relevant.  The loops are there simply to help modeller's intuition.

{\bf Phil asks: how much training was necessary before the volunteers
understood how to work the interface? Could this training be done with
no input from the \spl developers, in future? It would be good to
comment on how scalable this approach to modeling is.}

As implemented so far, \spl assumes that the lens is dominated by a
single galaxy.  Only one maximum in the arrival-time surface
is permitted, and it is taken to be the centre of the main lensing
galaxy.  The user can, however, mark additional minor galaxies: these
are modelled as point masses, the mass being fitted by the program along
with the rest of the mass distribution.

@@

Having sketched a spaghetti diagram, the user presses a button to
initiate the next stage.  \spl then translates the spaghetti diagram
into input for GLASS, and forwards this input.  The task of GLASS,
which runs server-side as it is compute-intensive, is to find a mass
distribution $\kappa(x,y)$ that exactly reproduces the given locations
of the maximum, minima and saddle points. This criterion by itself is
extremely under-determined --- there are infinitely many mass
distributions that will reproduce a given set of maxima, minima and
saddle points, but typically they (a)~produce lots of extra images,
and (b)~look very unlike galaxies.  Additional assumptions (a prior)
are necessary.  GLASS uses the following prior
\citep[cf.][]{1997MNRAS.292..148S,2008ApJ...679...17C}.
\begin{enumerate}
\item The mass distribution is built out of non-negative tiles of
  mass.  (Sometimes these tiles are called mass pixels, but we should
  emphasize that they are unrelated to image pixels, and are much
  larger.)
\item There is a notional lens center, say $(x_0,y_0)$ which is
  identified with the maximum of the arrival time.  The source can
  have an arbitrary offset with respect to the lens center.
\item The mass distribution must be centrally concentrated, in two
  respects.  First, the circularly averaged density must fall away
  like $$ \left[(x-x_0)^2+(y-y_0)^2\right]^{-1/2}$$ or more steeply.
  Second, the direction of increasing density at any $(x,y)$ can point
  at most $45^\circ$ away from $(x_0,y_0)$.
\item The lens must be symmetrical with respect to $180^\circ$ rotations
  about $(x_0,y_0)$.  This symmetry assumption can be relaxed if the
  user wishes.
\end{enumerate}
There are still infinitely many models that satisfy both data and
prior constraints, but now they are more credible as galaxy lenses.
It is then possible to generate an ensemble of models.  The sampling
technique used by GLASS is described in \citep{Lubini2012}.
Typically, ensembles of 200 models are used.  That is to say, what we
call a \spl model is really the mean of an ensemble of 200 models, and
its estimated uncertainty is the range covered by the whole ensemble.

@@

After the model ensemble has been generated, \spl post-processes it to
present results and diagnostics to the user for inspection. This takes
the form of three figures.
\begin{enumerate}
\item A synthetic image of the lensed features.
\item A contour map of the arrival-time surface $t(x,y)$.
\item A gray scale plus contour map of the mass distribution.
\end{enumerate}
The synthetic image generated by \spl assumes a simple
\phil{circularly symmetric} source \phil{with linearly decreasing
surface brightness profile}.  The user can change the contrast level
on the image, which (though it is not saved) amounts to adjusting the
\phil{size of the source}. These synthetic images are still very
crude, and \phil{not always} useful for assessing models.  The best
indicator, in practice, of whether the modelling was successful is
contour map of $t(x,y)$, with saddle-point contours highlighted.  It
is, in effect, the computer's refinement of the spaghetti diagram
input by the user.  If the arrival-time surface looks qualitatively
similar to the spaghetti diagram, that generally indicates a
successful model.  The mass distribution also provides indications;
successful models generally lead to smooth-looking mass distribution,
whereas an irregular or checkboard pattern in the mass map signals a
bad model.

After examining this feedback, the user can choose to \phil{save the
model to the \spl archive, at which point it is assigned a unique
URL}.  They can also modify their input and try again, or discard the
attempt altogether.  After archiving, there can be discussion among
modellers, through the \sw forum or by any other means, and revision
of the model.  \phil{This is achieved simply by sharing the model's
URL; following its hyperlink takes one to the \spl app, pre-loaded
with the correct data image and input spaghetti.} Any archived model
can be revised by any user: they can modify the spaghetti
configuration slightly or drastically, or change options like the size
of the mass tiles. Particularly interesting lens candidates lead to
trees of models in this way.  Discussion among modellers tends to
prune a model tree, focusing attention on the most interesting
models.
% PJM: sadly this Letter is no longer available, Julianne took it
% down. We can post the talk URL instead.
% \footnote{See ``Collaborative gravitational lens
% modelling\dots'' in {\tt http://letters.zooniverse.org} for an
% example.}
\footnote{See the ``collaborative modelling'' thread at 
{\tt http://talk.spacewarps.org} for an example.}
% TODO: fix url!

@@

Each of Figures \ref{fig:6941}--\ref{fig:7022} figures has the
following layout.
$$ \begin{matrix}
\hbox{marked-up image} \qquad &\hbox{model synthetic image} \\
t(x,y)                        &\hbox{model } t(x,y) \\
\kappa(x,y)                   &\hbox{model } \kappa(x,y)
\end{matrix} $$

@@

\noindent The marked-up image is a zoom of the lensed image on \sw,
marked up with one or more spaghetti contours; this is the modeller's
input to \spl.  The three panels on the right show the graphical
output returned to the modeller by \spl, listed as (i), (ii), (iii) in
\secref{SpaghettiLens}.  These derive from the mean of an ensemble of
200 models generated by \spl. The model synthetic image is too crude
to be useful, because there is no fitting for the source light
distribution, as already noted in \secref{diag}.  The quality of the
synthetic images can be improved by interpolation, but as this feature
was not available during the modelling challenge, we do not show the
improved versions here.

@@

To illustrate, \figref{kapenc_compare_faulty} compares the circularly
averaged mass profiles three different models of one particular lens;
two of the models are shown in Figures~\ref{fig:7025} and
\ref{fig:7022}. Each panel in \figref{kapenc_compare_faulty} shows the
mean $\kappa$ within a circle of given radius.  The red curve is the
correct profile for the sim.  The two blue curves are the minimal and
maximal mean enclosed $\kappa$ from the internal ensemble in \spl.
Radial locations of the images are marked, along with the image
parities.  The region between the blue curves is shaded between the
radii of the innermost and the outermost images: this is the
confidence region from the modelling.  The definition \eqref{eq:effRE}
for \ERf corresponds to crossing the dashed horizontal line at $10^0$:
the red curve crosses the dashed line at the actual Einstein radius
\ERf[,act]; the recovered Einstein radius \ERf[,rec] and its
uncertainty are given by the blue curves crossing the dashed line.  We
see that in all three panels, the blue curves are shallower than the
red curve and \ERf[,rec] is more than \ERf[,act], by more than the
model uncertainties.  Now, steeper mass profiles tend to give wider
image separations --- recall that the image separation for a circular
isothermal lens is 2\ERf, whereas for a point mass it is more
\citep[see, e.g.,][]{2002LNP...608....1C} --- so \ERf[,rec] being too
high is really a consequence of the \phil{GLASS} models being too shallow for
the sims.

@@

\Figref{ER_per_sim} shows that \ERf[,rec] of the models tend to be too
high.  \phil{However, this is entirely due to the GLASS model density
profiles being too shallow, as illustrated above. We can separate out
the performance of the \spl interface and its users by comparing their
results with the  Einstein radii of \spl models made by an expert
(PS). Discounting the models which were flagged by the volunteers as
poor, the mean Einstein radius overestimate was 25\%, with a 20\%
standard deviation (shown by the grey band in \Figref{ER_per_sim}). 
The expert models show a statistically indistinguishable bias of 20\%,
with standard deviation 8\% (the blue band in \Figref{ER_per_sim}).  }
One source of this systematic error is that it is difficult to center
of the lens accurately: an offset leads to a flatter mass profile for
the model compared to the simulation.

@@

\spl is an implementation of these ideas, enabling experienced but
non-professional lens enthusiasts to model newly-discovered lens
candidates from the \sw citizen-science platform.  The tests in this
paper indicate that such modelling would be both feasible and
scientifically interesting: \phil{given a suitable modelling tool, and
appropriate guidance, a small team of non-professional volunteers was
able to model a sample of 29 test lenses, and measure their Einstein
radii with comparable accuracy to a professional expert.}

There is, however, plenty of room for improvement:
\begin{enumerate}
\item \spl tends to overestimate the Einstein radius (evident from
  \Figref{ER_per_sim}), and it is also apparent that the model
  density profiles tend to be too shallow.  The likely explanation is
  that while the sims are steeply peaked at the centre, the pixellated
  mass model fixes a comparatively large area near the central at
  constant density.  One way to solve this problem would be to
  introduce smaller pixels in the central region, thus enabling a
  steeper centre. \phil{The use of simply parameterised, appropriately
  steeply-profiled models would also avoid this problem.}
\item Currently, \spl does not attempt to model the source shape; the
  user identifies the brightest points on the image, and these are
  taken as images of a point-like source, whose positions must be
  reproduced exactly. For generating a synthetic image, a conical
  source profile is assumed. Fitting for the source profile to
  optimize resemblance to the observed lensed image after the lens
  model has been generated, is algorithmically straightforward
  \citep[cf.][]{2003ApJ...590..673W,2006MNRAS.371..983S} and planned
  to be implemented.  This would alleviate another problem with \spl,
  which is that there is as yet no quantitative figure of merit for 
  any given model: assessment of each model is a judgment call based
  on the synthetic image, and on whether the mass distribution and the
  arrival-time surface show suspicious features.
\item Another limitation so far in \spl is that the lens is assumed to
  be dominated by one galaxy, which puts most galaxy-group lenses
  beyond the reach of the modeler. Since complicated group lenses are
  some of the most interesting candidates present, removing this
  limitation is most desirable.  From the users' point of view, it
  would mean that spaghetti contours with more than one maximum can be
  allowed.  For examples, see Figure 5c in \citep{2001ApJ...557..594R}
  and Figure 4b in \cite{2003ApJ...590...39K}.
\item At present, a single false-color composite is used as the data.
  An option could be added to use all available filters, individually
  or in combination, at the user wishes. 
\item As mentioned above, the option of revising an already-archived
  model is already available.  Desired now are tools for comparing
  different models of a given system, both visually and through
  different statistical measures.  As evidenced by a current
  collaborative modeling effort, a particularly interesting candidate
  can lead to an extended discussion and dozens of models, \phil{that
  in some way sample the high likelihood region of model parameter
  space}.
\item Better tutorial materials are also needed, and this would
  address some of the problem areas found in the modeling challenge.
  For example, we saw in \secref{tests.t1} that volunteers are most
  prone to making errors in two situations: when in identifying an arc-
  like structure while placing the points, and in identifying the
  correct ordering of the points in nearly-symmetric configurations.
\end{enumerate}

>>>>

\definecolor{purple}{RGB}{150,0,200}
\newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf #1}}

@@

We develop a method to enable collaborative modelling of gravitational
lenses, and lens candidates, by non-professional lens enthusiasts.  It
uses an existing free-form modelling program (GLASS), but
enables the input to this code to be provided in a novel way,
via a user-provided diagram that is essentially a sketch of an
arrival-time surface.

An implementation of this method, \spl is tested in a modelling
challenge using 29 simulated lenses drawn from the set created
for the \sw citizen-science lens search.  We find that
volunteers from this online community asserted the image
parities and time ordering consistently in some lenses, but made
errors in other lenses depending on the image morphology. While
errors in image parity and time ordering lead to  large errors in the
mass distribution, the enclosed mass was found to be more
robust: the model-derived Einstein radii found by the volunteers were
consistent with those produced by one of the professional team,
suggesting that given the appropriate tools, gravitational lens
modelling is a data analysis activity that can be crowd-sourced to
good effect. Shortcomings of the software implementation 
include (a)~lens profiles that tend to be systematically shallower
than the lenses (leading to systematic over-estimation of the
Einstein radius  by some $\sim25\%$, and (b)~lack of detailed
arc modelling.  Ideas for improvement are discussed.

@@

Imaging surveys now under way aim to increase the inventory of lenses
another ten or a hundred fold \citep[see
e.g.][]{Marshall2005,OguriMarshall2010}, with both automated and
visual search techniques proposed
\citep[e.g.][]{Marshall2009,More2012ApJ,Gavazzi2014}. For example, \sw
(Marshall et al, in prep, More et al, in prep) is a citizen science
project\footnote{\tt http://www.spacewarps.org} in which volunteers
are presented sky-survey images and invited to identify lens
candidates, by eye.  Simulated lenses are mixed in with the data,
both to help train volunteers on what to look for, and to provide
measures of the effectiveness of the search.  The motivation for \sw
is to enable volunteers, some of whom had previously serendipitously
identified lens candidates on earlier citizen-science surveys,
either to make discoveries missed in automatic searches by
software robots, or to perform the necessary inspection of an
automatically-generated sample, for quality control.  Robots can be
built to be good at detecting clean lensing system in uncrowded fields
with high signal-to-noise, but in more general test situations robots
miss lenses (low completeness) or contaminate the results with
non-lenses (low purity) \citep{Marshall2009}.

The encouraging early results from the first \sw lens search,
carried out on the  $\simeq172$ square degree Canada-France-Hawaii
Telescope Legacy Survey (CFHTLS) imaging by over 30,000 volunteers
(Marshall et al, in prep, More et al, in prep) prompted the question:
could the modelling of the lenses also be done by the volunteers?  If
so, modelling could help assess prioritise lens candidates at an early
stage, which would be very useful with new wide-field and sensitive
surveys, which will yield thousands of lens candidates. There are
several software tools for lensing modelling available, and work has
been done on generic interfaces \citep[e.g.][]{2014A&C.....5...28L}. 
Indeed, some early designs for \sw included a prototype lens modelling
tool \citep{2010AAS...21543527N}. Moreover, some \sw volunteers are
quite experienced from earlier projects, having spent a thousand hours
or more with data, and are very interested in more demanding
projects.  The interests of citizen-science communities are just
beginning to be studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but it
is clear that some volunteers welcome open-ended challenges, and
sometimes these have led to new scientific results: one example is the
discovery of an exceptional extra-solar planet
\citep{2013ApJ...768..127S}; another is the development of new
algorithms for protein folding \citep{Khatib22112011}.  All these are
grounds for optimism.  There is, however, a basic difficulty in strong
gravitational lensing. Lensed images do not look much like their
source, and still less do they resemble the lensing-mass
distribution.  To model a lens, one needs either to do a lot of random
guessing, or to have a good intuition for what works.

@@

In \secref{SpaghettiLens} we describe the \spl program, which
implements the above scheme.  \spl is an interface to and
extension of the GLASS framework for modelling lenses
\citep{2014arXiv1401.7990C}.  We will not go into software details in
this paper, but instead concentrate on lens modelling per se, but we
remark that \spl is designed to be friendly to the forum style of
citizen-science projects, and enable incremental collaborative
model refinement by different people, without sacrificing any of the
technical features of GLASS.

In \secref{mod_challenge} we describe a modelling challenge where a
diverse sample of 29 simulated lenses was modelled multiple times
by a small number of \sw volunteers using \spl.  The models
were then examined in two ways.  One was whether the spaghetti diagram
was correct.  The other was the recovery of the Einstein radius of the
lens.  In addition, we show some visual comparisons of the actual and
recovered lens shape and radial profile, and identify some areas to
improve. Profile and shape recovery with GLASS has been studied in
more detail in \citep{2014arXiv1401.7990C}.

@@

\spl is not part of \sw, but it makes use of the \sw infrastructure,
in particular, the image database and the discussion
forum.\footnote{\tt http://talk.spacewarps.org}  The forum is
essential for establishing contact between interested members of the
\sw community and the project science team, and then for enabling
collaboration between them. We were able to collaborate together on 
modelling objects from \sw in the usual style of medium-sized
astronomical collaborations, with video-conferencing and in-person
meetings where possible. Preliminary results were immediately
summarized on modelling threads on the forum, and anyone interested
was made welcome to join at any time.
% {\bf Phil asks: how much training was necessary before the volunteers
% understood how to work the interface? Could this training be done with
% no input from the \spl developers, in future? It would be good to
% comment on how scalable this approach to modeling is.}

% PJM: The following paragraph is not accurate: the manipulation of 
%  FITS images in the Tools viewer, for example, is done client side.
%  In future I very much hope to see lens modeling enabled in this
%  mode as well!
%
% Another aspect of \sw (and of the Zooniverse family of projects in
% general) that \spl adopts, is to keep all user interaction in a web
% browser, with intensive computing done on a server.  Several
% interesting design and implementation issues arise from these
% considerations, but we will not go into them in this paper,
% concentrating instead on aspects specific to lens modelling.
%

Modelling with \spl involves three stages, markup of the image,
followed by intensive numerical computation carried out on a
server in the background, followed by review
of diagnostics and possible discussion.  We now describe each of
these.

@@

The markup tool allows only valid lensing configurations to be
entered.  The user does not need to think explicitly about the image
parities (though the markup tool provides this information using
colour codes) or about time-ordering, or worry about the odd-image
theorem.  The exact placement of the loops in a spaghetti diagram has
no significance.  Only the hierarchy of which loop is inside which is
relevant.  The loops are there simply to help modeller's intuition.

As implemented so far, \spl assumes that the lens is dominated by a
single galaxy.  Accordingly, only one maximum in the arrival-time surface
is permitted, and it is taken to be the centre of the main lensing
galaxy.  The user can, however, mark additional minor galaxies: these
are modelled as point masses, the mass being fitted by the program along
with the rest of the mass distribution.

@@

Having sketched a spaghetti diagram, the user presses a button to
initiate the next stage.  \spl then translates the spaghetti diagram
into input for GLASS, and forwards this input.  The task of GLASS,
which runs server-side as it is compute-intensive, is to find a mass
distribution $\kappa(x,y)$ that exactly reproduces the given locations
of the maximum, minima and saddle points. This criterion by itself is
extremely under-determined --- there are infinitely many mass
distributions that will reproduce a given set of maxima, minima and
saddle points, but typically they (a)~produce lots of extra images,
and (b)~look very unlike galaxies.  Additional assumptions (a prior)
are necessary.  GLASS uses the following prior
\citep[cf.][]{1997MNRAS.292..148S,2008ApJ...679...17C}.
\begin{enumerate}
\item The mass distribution is built out of non-negative tiles of
  mass.  (Sometimes these tiles are called mass pixels, but we should
  emphasize that they are unrelated to image pixels, and are much
  larger.)
\item There is a notional lens centre, say $(x_0,y_0)$ which is
  identified with the maximum of the arrival time.  The source can
  have an arbitrary offset with respect to the lens centre.
\item The mass distribution must be centrally concentrated, in two
  respects.  First, the circularly averaged density must fall away
  like $$ \left[(x-x_0)^2+(y-y_0)^2\right]^{-1/2}$$ or more steeply.
  Second, the direction of increasing density at any $(x,y)$ can point
  at most $45^\circ$ away from $(x_0,y_0)$.
\item The lens must be symmetrical with respect to $180^\circ$ rotations
  about $(x_0,y_0)$.  This symmetry assumption can be relaxed if the
  user wishes.
\end{enumerate}
There are still infinitely many models that satisfy both data and
prior constraints, but now they are more credible as galaxy lenses.
It is then possible to generate an ensemble of models.  The sampling
technique used by GLASS is described in \citep{Lubini2012}.
Typically, ensembles of 200 models are used.  That is to say, what we
call a \spl model is really the mean of an ensemble of 200 models, and
its estimated uncertainty is the range covered by the whole ensemble.

@@

After the model ensemble has been generated, \spl post-processes it to
present results and diagnostics to the user for inspection. This takes
the form of three figures.
\begin{enumerate}
\item A synthetic image of the lensed features.
\item A contour map of the arrival-time surface $t(x,y)$.
\item A gray scale plus contour map of the mass distribution.
\end{enumerate}
The synthetic image generated by \spl assumes a simple
circularly symmetric source with linearly decreasing
surface brightness profile.  The user can change the contrast level
on the image, which (though it is not saved) amounts to adjusting the
size of the source. These synthetic images are still very
crude, and not always useful for assessing models.  The best
indicator, in practice, of whether the modelling was successful is
contour map of $t(x,y)$, with saddle-point contours highlighted.  It
is, in effect, the computer's refinement of the spaghetti diagram
input by the user.  If the arrival-time surface looks qualitatively
similar to the spaghetti diagram, that generally indicates a
successful model.  The mass distribution also provides indications;
successful models generally lead to smooth-looking mass distribution,
whereas an irregular or checkboard pattern in the mass map signals a
bad model.

After examining this feedback, the user can choose to save the
model to the \spl archive, at which point it is assigned a unique
URL.  They can also modify their input and try again, or discard the
attempt altogether.  After archiving, there can be discussion among
modellers, through the \sw forum or by any other means, and revision
of the model.  This is achieved simply by sharing the model's
URL; following its hyperlink takes one to the \spl app, pre-loaded
with the correct data image and input spaghetti. Any archived model
can be revised by any user: they can modify the spaghetti
configuration slightly or drastically, or change options like the size
of the mass tiles. Particularly interesting lens candidates lead to
trees of models in this way.  Discussion among modellers tends to
prune a model tree, focusing attention on the most interesting
models.
% PJM: sadly this Letter is no longer available, Julianne took it
% down. We can post the talk URL instead.
% \footnote{See ``Collaborative gravitational lens
% modelling\dots'' in {\tt http://letters.zooniverse.org} for an
% example.}
\footnote{See the ``collaborative modelling'' thread at 
{\tt http://talk.spacewarps.org} for an example.}
% TODO: fix url!

@@

Each of Figures \ref{fig:6941}--\ref{fig:7022} figures has the
following layout.
$$ \begin{matrix}
\hbox{marked-up image} \qquad &\hbox{model synthetic image} \\
t(x,y)                        &\hbox{model } t(x,y) \\
\kappa(x,y)                   &\hbox{model } \kappa(x,y)
\end{matrix} $$
The marked-up image is a zoom of the lensed image on \sw,
marked up with one or more spaghetti contours; this is the modeller's
input to \spl.  The three panels on the right show the graphical
output returned to the modeller by \spl, listed as (i), (ii), (iii) in
\secref{SpaghettiLens}.  These derive from the mean of an ensemble of
200 models generated by \spl. The model synthetic image is too crude
to be useful, because there is no fitting for the source light
distribution, as already noted in \secref{diag}.  The quality of the
synthetic images can be improved by interpolation, but as this feature
was not available during the modelling challenge, we do not show the
improved versions here.

@@

@@

To illustrate, \figref{kapenc_compare_faulty} compares the circularly
averaged mass profiles three different models of one particular lens;
two of the models are shown in Figures~\ref{fig:7025} and
\ref{fig:7022}. Each panel in \figref{kapenc_compare_faulty} shows the
mean $\kappa$ within a circle of given radius.  The red curve is the
correct profile for the sim.  The two blue curves are the minimal and
maximal mean enclosed $\kappa$ from the internal ensemble in \spl.
Radial locations of the images are marked, along with the image
parities.  The region between the blue curves is shaded between the
radii of the innermost and the outermost images: this is the
confidence region from the modelling.  The definition \eqref{eq:effRE}
for \ERf corresponds to crossing the dashed horizontal line at $10^0$:
the red curve crosses the dashed line at the actual Einstein radius
\ERf[,act]; the recovered Einstein radius \ERf[,rec] and its
uncertainty are given by the blue curves crossing the dashed line.  We
see that in all three panels, the blue curves are shallower than the
red curve and \ERf[,rec] is more than \ERf[,act], by more than the
model uncertainties.  Now, steeper mass profiles tend to give wider
image separations --- recall that the image separation for a circular
isothermal lens is 2\ERf, whereas for a point mass it is more
\citep[see, e.g.,][]{2002LNP...608....1C} --- so \ERf[,rec] being too
high is really a consequence of the GLASS models being too shallow for
the sims.

@@

\Figref{ER_per_sim} shows that \ERf[,rec] of the models tend to be too
high.  However, this is entirely due to the GLASS model density
profiles being too shallow, as illustrated above. We can separate out
the performance of the \spl interface and its users by comparing their
results with the  Einstein radii of \spl models made by an expert
(PS). Discounting the models which were flagged by the volunteers as
poor, the mean Einstein radius overestimate was 25\%, with a 20\%
standard deviation (shown by the grey band in \Figref{ER_per_sim}). 
The expert models show a statistically indistinguishable bias of 20\%,
with standard deviation 8\% (the blue band in \Figref{ER_per_sim}).
One source of this systematic error is that it is difficult to centre
of the lens accurately: an offset leads to a flatter mass profile for
the model compared to the simulation.

@@

\spl is an implementation of these ideas, enabling experienced but
non-professional lens enthusiasts to model newly-discovered lens
candidates from the \sw citizen-science platform.  The tests in this
paper indicate that such modelling would be both feasible and
scientifically interesting: given a suitable modelling tool, and
appropriate guidance, a small team of non-professional volunteers was
able to model a sample of 29 test lenses, and measure their Einstein
radii with comparable accuracy to a professional expert.

There is, however, plenty of room for improvement:
\begin{enumerate}
\item \spl tends to overestimate the Einstein radius (evident from
  \Figref{ER_per_sim}), and it is also apparent that the model
  density profiles tend to be too shallow.  The likely explanation is
  that while the sims are steeply peaked at the centre, the pixellated
  mass model fixes a comparatively large area near the central at
  constant density.  One way to solve this problem would be to
  introduce smaller pixels in the central region, thus enabling a
  steeper centre. The use of simply parameterised, appropriately
  steeply-profiled models would also avoid this problem.
\item Currently, \spl does not attempt to model the source shape; the
  user identifies the brightest points on the image, and these are
  taken as images of a point-like source, whose positions must be
  reproduced exactly. For generating a synthetic image, a conical
  source profile is assumed. Fitting for the source profile to
  optimize resemblance to the observed lensed image after the lens
  model has been generated, is algorithmically straightforward
  \citep[cf.][]{2003ApJ...590..673W,2006MNRAS.371..983S} and planned
  to be implemented.  This would alleviate another problem with \spl,
  which is that there is as yet no quantitative figure of merit for 
  any given model: assessment of each model is a judgment call based
  on the synthetic image, and on whether the mass distribution and the
  arrival-time surface show suspicious features.
\item Another limitation so far in \spl is that the lens is assumed to
  be dominated by one galaxy, which puts most galaxy-group lenses
  beyond the reach of the modeler. Since complicated group lenses are
  some of the most interesting candidates present, removing this
  limitation is most desirable.  From the users' point of view, it
  would mean that spaghetti contours with more than one maximum can be
  allowed.  For examples, see Figure 5c in \citep{2001ApJ...557..594R}
  and Figure 4b in \cite{2003ApJ...590...39K}.
\item At present, a single false-color composite is used as the data.
  An option could be added to use all available filters, individually
  or in combination, at the user wishes. 
\item As mentioned above, the option of revising an already-archived
  model is already available.  Desired now are tools for comparing
  different models of a given system, both visually and through
  different statistical measures.  As evidenced by a current
  collaborative modeling effort, a particularly interesting candidate
  can lead to an extended discussion and dozens of models, that
  in some way sample the high likelihood region of model parameter
  space.
\item Better tutorial materials are also needed, and this would
  address some of the problem areas found in the modeling challenge.
  For example, we saw in \secref{tests.t1} that volunteers are most
  prone to making errors in two situations: when in identifying an arc-
  like structure while placing the points, and in identifying the
  correct ordering of the points in nearly-symmetric configurations.
\end{enumerate}
